---
title: "HW0"
author: "Rachel Sussman"
format: html
execute:
  echo: false
---

## Exercises

(@) **Write a Quarto document that defines variables $a=1, b=3, c=2$
and print out the solutions to $f(x) = ax^2+bx+c=0$. Do not report complex solutions, only real numbers.**

```{r}
a=1
b=3
c=2

if(b^2 == 4*a*c){
  p1 = paste0("there is one solution. The solution is x=", (-b + sqrt(b^2 - 4*a*c))/(2*a), ".")
  p2 = paste0("As indicated by the red dashed line, there is one point where the quadratic curve intersects with the line y=0: at x=", (-b + sqrt(b^2 - 4*a*c))/(2*a), ".")
}else if(b^2 > 4*a*c){
  p1 = paste0("there are two solutions. The solutions are x=", (-b + sqrt(b^2 - 4*a*c))/(2*a), " and x=", (-b - sqrt(b^2 - 4*a*c))/(2*a), ".")
  p2 = paste0("As indicated by the red dashed lines, there are two points where the quadratic curve intersects with the line y=0: one at x=", (-b + sqrt(b^2 - 4*a*c))/(2*a), " and the other at x=", (-b - sqrt(b^2 - 4*a*c))/(2*a), ".")
}else if(b^2 < 4*a*c){
  p1 = paste0("there are no real solutions.")
  p2 = paste0("As you can see, the quadratic curve does not intersect with the line y=0; hence, there are no solutions.")
}
```

The quadratic formula can be applied to solve this equation: $$\frac{-b \pm \sqrt{b^{2}-4ac}}{2a}$$ Specifically, the variables *a*, *b*, and *c* are entered into the above formula to find the solutions.

A quadratic equation can have 0, 1, or 2 solutions based on its *discriminant*: $b^{2}-4ac$. A quadratic equation has no real solutions when the discriminant is negative, one solution when the discriminant equals zero, and two solutions when the discriminant is positive. 

My code checks the discriminant using an if-else statement where the conditionals are the possibilities for the discriminant listed above (negative, zero, or positive) and the instructions are to evaluate the quadratic formula (if the discriminant $\geq$ zero), print out the number and value(s) of the solution(s), and plot and explain the outcome.

For the quadratic equation $`r a`x^{2} + `r b`x + `r c` = 0$, `r p1`

(@) **Include a graph of $f(x)$ versus $x$ for $x \in (-5,5)$.**

```{r}
x=seq(-5, 5, length.out = 300)
plot(x, a*x^2 + b*x + c, type = "l")
abline(h = 0, lty = 2)

if(b^2 == 4*a*c){
  abline(v = (-b + sqrt(b^2 - 4*a*c))/(2*a), col = "red", lty=2)
}else if(b^2 > 4*a*c){
    abline(v = (-b + sqrt(b^2 - 4*a*c))/(2*a), col = "red", lty=2)
    abline(v = (-b - sqrt(b^2 - 4*a*c))/(2*a), col = "red", lty=2)
}
```
Here is a graph of $f(x)$ versus $x$ for $x \in (-5,5)$. `r p2`

<!-- (@) **Generate a PDF report using knitr. Do not show the R code, only the solutions and explanations of what the reader is seeing.** -->

<!-- In the terminal, enter `quarto render HW0.qmd --to pdf`. Alternatively, in the RStudio console, enter `rmarkdown::render("HW0.qmd", "pdf_document")`. To hide the R code, type `execute: echo: false` in the Quarto file header or  `echo = FALSE` in each bracket that initializes a code chunk. -->

<!-- (@) **Erase the PDF report and reproduce it but this time using $a=1, b=2, c=5$.** -->

<!-- To erase the PDF report, enter `rm HW0.pdf` in the terminal. Then, to reproduce the report using $a=1, b=2$, and $c=5$, assign 1, 2, and 5 to variables a, b, and c respectively on lines 15-17 of this Quarto file, then enter `quarto render HW0.qmd --to pdf` in the terminal or enter `rmarkdown::render("HW0.qmd", "pdf_document")` in the RStudio console. -->

<!-- (@) **Erase the PDF report and reproduce it but this time using $a=1, b=3, c=2$.** -->

<!-- To erase the PDF report, enter `rm HW0.pdf` in the terminal. Then, reproduce the report using $a=1, b=3$, and $c=2$, assign 1, 3, and 2 to variables a, b, and c respectively on lines 15-17 on this Quarto file, then enter `quarto render HW0.qmd --to pdf` in the terminal or enter `rmarkdown::render("HW0.qmd", "pdf_document")` in the RStudio console. -->

<!-- (@) **Create an HTML page with the results for this last set of values, but this time showing the code.** -->

<!-- To show the R code, either type `execute: echo: true` (or delete `execute: echo: false` if applicable) in the header or type `echo = TRUE` (or delete `echo = FALSE` if applicable) in each bracket that initializes a code chunk. Then, in the terminal, enter `quarto render HW0.qmd --to html`, or in the RStudio console, enter `rmarkdown::render("HW0.qmd", "html_document")`. -->
